% !TeX root = ../main.tex
% Add the above to each chapter to make compiling the PDF easier in some editors.

\chapter{Experiment setup and Methodology}\label{chapter:experiment_setup}
In this chapter we will discuss about our experiment setup and methodology for training our neural network models. Our goal is to preprocess the data from our simulation software in a format that we can feed it to the neural network models, then train the models. The data we got from our simulation is not suitable for direct feeding to the neural network because different models have different structure for the input of the data. We will also discuss the mythology behind the formatting of the data and our experiment setup and also the architectures of our Convolutional Neural Network (CNN).


\section{Data Preprocessing}

Our data preprocessing mainly contain 2 steps. First align the IMU data points (number of rows) according to the data points of our Ground Truth. Then fine tune the data in order to make it suitable for calculation of the CNN.

\subsection{Aligning the data rows}

As mention before in the section ooooo we took our IMU data and Ground Truth data in different files. For each IMU we have 2 files for accelerometer data from the accelerometer sensor and angular velocity data from the gyroscope. The Ground Truth data has one file for each simulation run.  The first thing is to check is the now of rows we got from each file. Our goal is to ensure that the count of data rows in the acceleration and angular velocity files was identical to the count of data rows in the GT file to ensure that they were fed smoothly into the model and keep the data sync with timestamps. Also, we wanted to ensure there are no “NaN” values in the data set to ensure that our model could learn optimally.

Our target was to get 100 thousand data rows. For that, we needed to run the simulation for more than an hour. We were expecting to get the same number of data rows from each file but unfortunately, that was not the case. We have a similar number of data rows form each IMU accelerometer and gyroscope, but the GT data rows were a bit less than the number of data rows from the IMU. For example, if we get ‘n’ number of data rows from each IMU file we were getting ‘n-x’ number of data rows from the GT.

Our assumption behind this problem is that when we stop the simulation by clicking the stop button The whole simulation doesn’t stop immediately. The plane stop moving but the Sensor still keeps generation some data on the other hand the GT stop generation data the moment we click stop because the GT script is a nested script, unlike the IMU script.
To overcome this problem, we used a simple solution.  We simply removed x data points from the data files of the accelerometer and the gyroscope to make all data files compatible with the same data point count. We measured the value of x is approximately 0.04\% of all data points which is ignorable.










































Citation test~\parencite{latex}.

\subsection{Subsection}

See~\autoref{tab:sample}, \autoref{fig:sample-drawing}, \autoref{fig:sample-plot}, \autoref{fig:sample-listing}.

\begin{table}[htpb]
  \caption[Example table]{An example for a simple table.}\label{tab:sample}
  \centering
  \begin{tabular}{l l l l}
    \toprule
      A & B & C & D \\
    \midrule
      1 & 2 & 1 & 2 \\
      2 & 3 & 2 & 3 \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{figure}[htpb]
  \centering
  % This should probably go into a file in figures/
  \begin{tikzpicture}[node distance=3cm]
    \node (R0) {$R_1$};
    \node (R1) [right of=R0] {$R_2$};
    \node (R2) [below of=R1] {$R_4$};
    \node (R3) [below of=R0] {$R_3$};
    \node (R4) [right of=R1] {$R_5$};

    \path[every node]
      (R0) edge (R1)
      (R0) edge (R3)
      (R3) edge (R2)
      (R2) edge (R1)
      (R1) edge (R4);
  \end{tikzpicture}
  \caption[Example drawing]{An example for a simple drawing.}\label{fig:sample-drawing}
\end{figure}

\begin{figure}[htpb]
  \centering

  \pgfplotstableset{col sep=&, row sep=\\}
  % This should probably go into a file in data/
  \pgfplotstableread{
    a & b    \\
    1 & 1000 \\
    2 & 1500 \\
    3 & 1600 \\
  }\exampleA
  \pgfplotstableread{
    a & b    \\
    1 & 1200 \\
    2 & 800 \\
    3 & 1400 \\
  }\exampleB
  % This should probably go into a file in figures/
  \begin{tikzpicture}
    \begin{axis}[
        ymin=0,
        legend style={legend pos=south east},
        grid,
        thick,
        ylabel=Y,
        xlabel=X
      ]
      \addplot table[x=a, y=b]{\exampleA};
      \addlegendentry{Example A};
      \addplot table[x=a, y=b]{\exampleB};
      \addlegendentry{Example B};
    \end{axis}
  \end{tikzpicture}
  \caption[Example plot]{An example for a simple plot.}\label{fig:sample-plot}
\end{figure}

\begin{figure}[htpb]
  \centering
  \begin{tabular}{c}
  \begin{lstlisting}[language=SQL]
    SELECT * FROM tbl WHERE tbl.str = "str"
  \end{lstlisting}
  \end{tabular}
  \caption[Example listing]{An example for a source code listing.}\label{fig:sample-listing}
\end{figure}
